opponent modelling

### 2020-10-30 comp_model_1019.Rmd
draft3

#### New work

## Fitting simple Bayesian learning model with flat prior

## Fit a bayesian model to observed human play. Prior is vector of possible level-k opponent strategies.
# # Function to produce best reponse to ai_action 
# best_response <- function(game,ai_action){
#   if (game == "rps"){
#     if (ai_action == "R") {return("P")}
#     else if (ai_action == "P") {return("S")}
#     else if (ai_action == "S") {return("R")}
#   } else if (game == "fwg"){ 
#     if (ai_action == "F") {return("W")}
#     else if (ai_action == "W") {return("G")}
#     else if (ai_action == "G") {return("F")}
#   } else if (game == "numbers") {
#     if (ai_action == "1") {return("2")}
#     else if (ai_action == "2") {return("3")}
#     else if (ai_action == "3") {return("4")}
#     else if (ai_action == "4") {return("5")}
#     else if (ai_action == "5") {return("1")}
# exp1_ToM <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {
#   # dat = data subset for one participant
#   # opp_strategy_vec = vector of possible opponent strtegies, model assumes humans restrict opp strategy space to vector   c("level0","level1","level2")
#   # return = -2logLik 
#   # or "likelihood_by_trial" vector  (for later plotting, etc)
#   # opp_mod_transfer is a boolean. True if distribution of opponent strategies is kept across games. False otherwise. 
#   # theta = parameter, probability computer will play its best response to what it thinks human is. Truth = 90%. in [0,1].
#   # eps: parameter controlling noise in human choice. Probability human will DEVIATE from its best response. In [0,1].
#   theta <- par[1]
#   eps <- par[2]
#   num_strat = length(opp_strategy_vec) 
#   # Load tables predicting oppponent move
#   rps_predict_opp <- read.csv("rps_predict_opp.csv")
#   fwg_predict_opp <- read.csv("fwg_predict_opp.csv")
#   numbers_predict_opp <- read.csv("numbers_predict_opp.csv")
#   #lik_opp is a vector that holds and update probability distribution of opp strategy given actions.
#   lik_opp <- matrix(0.0,num_strat, nrow(data))
#   # Initiate prior vector as uniform
#   opp_prior_vec <-rep(1/num_strat, num_strat)
#   # br_hum is vector that stores best responses of human to actions taken by each level-k ai opponent at the round.
#   br_hum <- rep(NA,num_strat)
#     pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)
#       pred_opp <- NA
#       ## KEY CODE FOR LEARNING TRANSFER !! (if no OM transfer, reset prior to uniform)
#       if (!(opp_mod_transfer)) {
#         opp_prior_vec <- rep(1/num_strat, num_strat)
#         lik_opp[,t] <- opp_prior_vec
#     # Rounds after first
#       # get prediction of opponent action from CSV files
#       for (strategy in opp_strategy_vec) {
#         k = match(strategy,opp_strategy_vec)
#         pred_opp <- as.character(filter(pred_file_opp, pred_file_opp$human_previous == as.character(data[t-1,"h_action"]) & pred_file_opp$computer_previous == as.character(data[t-1,"a_action"]))[[strategy]])
#         # Given opponent action predicted, what would be human action best response for each opp strat
#         br_hum[[k]] <- best_response(t_game,as.character(pred_opp))
#         # Multiply prior by likelihood of observation to get posterior. Done here ot take advantage of For loop. 
#         if(as.character(data[t,"a_action"]) == pred_opp) {
#           lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
#           lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
#       # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
#       indices <- which(br_hum %in% as.character(data[t,"h_action"]))
#       # THEN likelihood of current human action is just prior on opponent vec (haven't updated priors with curr opp act yet)
#       # if action is not predicted by any level-k OM, assume human chooses randomly with prob eps
#       if(length(indices) == 0){
#         lik_hum[t] <- eps/nopts
#       # if same br action for multiple possible ai opponents, add probabilities 
#         lik_hum[t] <- sum(opp_prior_vec[indices])*(1-eps) + eps/nopts
#       # Standardising the probabilities and updating prior
#       lik_opp[,t] <- lik_opp[,t]/sum(lik_opp[,t])
#       opp_prior_vec <- lik_opp[,t]
#     # for debugging uncomment the line below
#     #cat(as.character(pred_opp),as.character(data[t,"a_action"]),lik_opp[,t],"\n")
# # data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# # exp1_ToM(c(0.9,0.99),data,c("level0","level1","level2"),"-2loglik",FALSE)
# ```
# ```{r}
# # rps_predict_opp <- read.csv("rps_predict_opp.csv")
# # red_opp <- filter(rps_predict_opp, rps_predict_opp$human_previous == "R" & rps_predict_opp$computer_previous == "S")
# # 
# # as.character(red_opp[["level0"]])
# ```
# # Fitting a bayesian model WITH BETWEEN GAME TRANSFER of opp strat distribution between games 
# ```{r, cache=TRUE}
# exp1_ToM_TR <- list()
# for(id in unique(dat$human_id)) {
#   exp1_ToM_TR[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   # exp1_ToM_TR[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0,0), upper = c(1,1), data=tdat, opp_strategy_vec = c("level0","level1","level2") ,return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE, parallelType=1,parVar = c("best_response")))
#   exp1_ToM_TR[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
# save(exp1_ToM_TR,file="exp1_ToM_TR.RData")
# ```
# # Fitting a bayesian model with NO BETWEEN GAME TRANSFER of opp strat distribution between games 
# ```{r, cache =TRUE}
# exp1_ToM_NT <- list()
# for(id in unique(dat$human_id)) {
#   exp1_ToM_NT[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   # exp1_ToM_NT[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0.0,0.0), upper = c(1.0,1.0), data=tdat, opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE, parallelType=0,parVar = c("best_response")))
#   exp1_ToM_NT[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
# save(exp1_ToM_NT,file="exp1_ToM_NT.RData")
# ```
# # Summarise results in dataframe for both transfer and no transfer models 
# ```{r}
# Nbayes_par_results <- data.frame()
# for(id in unique(dat$human_id)) {
# Nbayes_par_results <- rbind(Nbayes_par_results,
#                        data.frame(
#                          "id" = id,
#                          "condition" = dat[dat$human_id==id,"condition"][1],
#                          # BICs 
#                          "BIC_transfer" = exp1_ToM_TR[[id]]$value + 2*log(150),
#                          "BIC_no_transfer" = exp1_ToM_NT[[id]]$value+ 2*log(150),
#                          # Theta is the parameter governing AI stochasticity. Truth is 0.9
#                          "theta_transfer" = exp1_ToM_TR[[id]]$par[1],
#                          "theta_no_transfer" = exp1_ToM_NT[[id]]$par[1],
#                          # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
#                          "eps_transfer" = exp1_ToM_TR[[id]]$par[2],
#                          "eps_no_transfer" = exp1_ToM_NT[[id]]$par[2]))
# #Plotting the theta/eps scatter for all participants...
# plot(Nbayes_par_results$theta_transfer,Nbayes_par_results$eps_transfer, xlab="Theta", ylab= "Epsilone",col = ifelse(as.character(Nbayes_par_results$condition) == "Level2",'red','green'))
# # 
# write.csv(Nbayes_par_results,file="Nbayes_par_results.csv",row.names = FALSE)

## Q-learning (Basic)
```{r}
Q_learn <- function(par,data,return_value=c("-2loglik","likelihood_by_trial")){
  # actions
  # beta= inverse temperature parameter in softmax choice function
  # lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]
  Q_vals_RPS = matrix(0.0,3)
  names(Q_vals_RPS) <- c("R","P","S")
  Q_vals_FWG = matrix(0.0,3)
  names(Q_vals_FWG) <- c("F","W","G")
  Q_vals_NUM = matrix(0.0,5)
  names(Q_vals_NUM) <- c("1","2","3","4","5")

@@

      # Assume human chooses action probabilistically using softmax on Q values
      probs <- exp(Q_vals[curr_state,]/beta)/sum(exp(Q_vals[curr_state,]/beta))
      # Get actual human next action and compute likelihood
      act_index <- match(h_act, colnames(Q_vals))

      lik_hum[t] <- probs[[act_index]]

@@

```{r}
# data34 <- subset(dat,human_id == "QSuzBXpbyRc370HsAACW")
# Q_learn_states(c(5, 0.5),data34 ,return_value = "likelihood_by_trial",gamma = 0 )

@@

##  Self-Tuning EWA with prior round play as state (Camerer & Ho 2007) 
```{r}

EWA_states <- function(par,data,return_value){

@@

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
      state_vec <- switch(as.character(t_game), rps=states_RPS, fwg = states_FWG, numbers = states_NUM)
      reward <- switch(as.character(t_game), rps=reward_RPS, fwg = reward_FWG, numbers = reward_NUM)
      game_data <- subset(data, game == t_game)
      # first round is uniform prediction
      lik_hum[t] <- 1/nopts
      N <- 1.0

      # Randomly select prev_state and actions for first round
      curr_state <- sample(state_vec, size = 1)
      h_act_prev <- sample(colnames(Att), size =1)
      a_act_prev <- sample(colnames(Att), size =1)
      #reward <- 0

      indx <- data[t,"round"]
      state_indx <- match(curr_state,state_vec)
      #cat("this is the current state index", state_indx, "\n")

      # Get reward and past human action
      h_act_prev <- as.character(game_data[indx-1,"h_action"])
      a_act_prev <- as.character(game_data[indx-1,"a_action"])
      new_state <- paste0(h_act_prev,a_act_prev)

      # Estimate phi(t)
      R_t <- as.numeric(colnames(Att) == a_act_prev)
      #cat("This is R(t)",R_t,"\n")

      H_t <- (H_t*(indx-2) + R_t)/(indx-1)
      #cat("this is H(t)",H_t,"\n")

@@

      for (i in 1:nopts) {
        action <- as.character(colnames(Att)[i])

        #cat("this is current indexed action",action,"\n")
        #cat("this is previous human action",h_act_prev,"\n")

        # Attraction vector update rule
        Att[state_indx,i] <- (Phi_t*N*Att[state_indx,i]  + ( delta_t[[i]] + (1- delta_t[[i]])*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (Phi_t*N + 1)
      }

@@

      # Assume human chooses action probabilistically using softmax on Attraction values
      probs <- exp(lambda*Att[new_state,])/sum(exp(lambda*Att[new_state,]))

      # Get actual human action and compute likelihood
      h_act <- as.character(game_data[indx,"h_action"])
      act_index <- match(h_act, colnames(Att))
      lik_hum[t] <- probs[[act_index]]

      curr_state <- new_state

      # if (indx == 50) {
      #   cat("this is the attraction matrix at last round of", "\n")
      #   print(Att)
      # }
    }
  }

@@

```{r, cache=TRUE}
# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# EWA_states(0.1, data, "-2loglik")
# optim(10.0,fn=EWA_states,gr = NULL, data=data,"-2loglik", lower = 0.0, upper = 200, method="L-BFGS-B")

EWA_states_modelling <- list()
for(id in unique(dat$human_id)) {
  EWA_states_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
#   EWA_states_modelling[[id]] <- optim(10.0,fn=EWA_states,gr = NULL, data=tdat,"-2loglik", lower = 0.0, upper = 200, method="L-BFGS-B")
# }
  EWA_states_modelling[[id]] <- DEoptim(fn=EWA_states, lower = 0.0, upper = 100.0, data=tdat, "-2loglik", control=list(trace = FALSE,parallelType=1))
}

save(EWA_states_modelling,file="EWA_states_modelling.RData")

@@

## Putting results together 
```{r}
# load("QL_modelling.RData")
# load("QL_states_modelling.RData")
# load("EWA_modelling.RData")
# load("EWA_self_modelling.RData")
# load("MBM_modelling.RData")
# load("EWA_states_modelling.RData")

@@

                         # Self-Tuning EWA (only 1 parameter)
                          # Parametric EWA BIC
                         "EWA_self_2LL" = EWA_self_modelling[[id]]$optim$bestval,
                         "EWA_self_BIC" = EWA_self_modelling[[id]]$optim$bestval + 1*log(150),
                         #Lambda is a parameter of the softmax choice function (inverse Temperature)
                         "EWA_self_Lambda" = EWA_self_modelling[[id]]$optim$bestmem[1],

                         # EWA with states

                         "EWA_states_BIC" = EWA_states_modelling[[id]]$optim$bestval + 1*log(150),
                         "EWA_states_Lambda" = EWA_states_modelling[[id]]$optim$bestmem[1]

@@

comp_BICs <- All_results[c("ID","Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC","EWA_states_BIC")]

@@

Table_results <- table(All_results[, "condition"],c("random","ToM_BT","ToM_NBT","QL", "QL_states","EWA","S_EWA","EWA_states")[apply(All_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC","EWA_states_BIC")],1,which.min)])

@@

exp1_model_comp <- data.frame()
for(id in unique(dat$human_id)) {
  tdat <- subset(dat,human_id == id)
  tot_score <- sum(tdat$score)
  tot_time <- sum(tdat$human_rt)
  #early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg"))
  early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg" | game =="numbers") )
  tr_score <- sum(early_dat$score)
  id_results <- subset(All_results, ID == id)
  min_BIC <- apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC")],1,min)

  best_model <- c("random","ToM_Game_Tr","ToM_No_Tr","QL", "QL_states","EWA","S_EWA", "EWA_states")[apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC","EWA_states_BIC")],1,which.min)]
  # 
  exp1_model_comp <- rbind(exp1_model_comp,
                       data.frame(
                         "human_id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         "Early_game_score" = tr_score,
                         "Total_score" = tot_score,
                         "Best_model" = best_model,
                         "Total_time" = sum(tdat$human_rt),
                         "TR_minus_NT_BIC" = id_results[,"Bayes_Tr_BIC"] - id_results[,"Bayes_No_Tr_BIC"],
                         "Rand_minus_best_BIC" =  id_results[,"Random_BIC"] - min_BIC

@@

>>>>

#### New work 
## Q-learning (Basic)
```{r}
Q_learn <- function(par,data,return_value=c("-2loglik","likelihood_by_trial")){
  # actions
  # beta= inverse temperature parameter in softmax choice function
  # lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]
  Q_vals_RPS = matrix(0.0,3)
  names(Q_vals_RPS) <- c("R","P","S")
  Q_vals_FWG = matrix(0.0,3)
  names(Q_vals_FWG) <- c("F","W","G")
  Q_vals_NUM = matrix(0.0,5)
  names(Q_vals_NUM) <- c("1","2","3","4","5")

@@

      # Assume human chooses action probabilistically using softmax on Q values
      probs <- exp(Q_vals[curr_state,]/beta)/sum(exp(Q_vals[curr_state,]/beta))

      # Get actual human next action and compute likelihood
      act_index <- match(h_act, colnames(Q_vals))
      lik_hum[t] <- probs[[act_index]]

@@

```{r}
data34 <- subset(dat,human_id == "QSuzBXpbyRc370HsAACW")

# Q_learn_states(c(5, 0.5),data34 ,return_value = "likelihood_by_trial",gamma = 0 )

@@

##  First attempt : Self-Tuning EWA with prior round play as state (Camerer & Ho 2007) 
# Initiates N(0) = 1 as in Camerer and Ho 1997 paper. 
# EWA_states <- function(par,data,return_value){
#   lambda <- par[1]
#   #Define matrix of state spaces for each game 
#   G1 <- expand.grid(c("R", "P", "S"),c("R", "P", "S"))
#   states_RPS <- paste0(G1$Var1,G1$Var2)
#   G2 <- expand.grid(c("F", "W", "G"), c("F", "W", "G"))
#   states_FWG <- paste0(G2$Var1,G2$Var2)
#   G3 <- expand.grid(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
#   states_NUM <- paste0(G3$Var1,G3$Var2)
#   A_RPS = matrix(-0.5,9,3)
#   dimnames(A_RPS) = list(states_RPS, c("R", "P", "S"))
#   A_FWG = matrix(-0.5,9,3)
#   dimnames(A_FWG) = list(states_FWG, c("F", "W", "G"))
#   A_NUM = matrix(-0.5,25,5)
#   dimnames(A_NUM) = list(states_NUM, c("1", "2", "3", "4", "5"))
#   # # Define reward matrices from the perspective of the row player (human in our case)
#   reward_RPS <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
#   dimnames(reward_RPS) = list(c("R", "P", "S"), c("R", "P", "S"))
#   reward_FWG <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
#   dimnames(reward_FWG) = list(c("F", "W", "G"), c("F", "W", "G"))
#   reward_NUM <- t(matrix(c(0,-1,0,0,1,1,0,-1,0,0,0,1,0,-1,0,0,0,1,0,-1,-1,0,0,1,0),nrow=5, ncol=5))
#   dimnames(reward_NUM) = list(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
#   # Initiate likelihood by trial vector 
#       Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
#       state_vec <- switch(as.character(t_game), rps=states_RPS, fwg = states_FWG, numbers = states_NUM)
#       reward <- switch(as.character(t_game), rps=reward_RPS, fwg = reward_FWG, numbers = reward_NUM)
#       game_data <- subset(data, game == t_game)
#       N <- 1.0
#       R_t <- rep(0.0,nopts)
#       H_t <- R_t
#       # Randomly select prev_state and actions for first round
#       curr_state <- sample(state_vec, size = 1)
#       h_act_prev <- sample(colnames(Att), size =1)
#       a_act_prev <- sample(colnames(Att), size =1)
#       #reward <- 0 
#       indx <- data[t,"round"]
#       state_indx <- match(curr_state,state_vec)
#       #cat("this is the current state index", state_indx, "\n")
#       # Get reward and past human action
#       h_act_prev <- as.character(game_data[indx-1,"h_action"])
#       a_act_prev <- as.character(game_data[indx-1,"a_action"])
#       new_state <- paste0(h_act_prev,a_act_prev)
#       # Estimate R(t) (recent history) then H(t) (history) then phi(t) (change detector)
#       R_t <- as.numeric(colnames(Att) == a_act_prev)
#       #cat("This is R(t)",R_t,"\n")
#       H_t <- (H_t*(indx-2) + R_t)/(indx-1)
#       #cat("this is H(t)",H_t,"\n")
#       Phi_t <- 1 - 0.5*sum((H_t-R_t)^2)
#       #cat("This is Phit(t)",Phi_t,"\t")
#       # Estimate vector Delta(t)
#       delta_t <- as.numeric(reward[,a_act_prev] >= as.numeric(game_data[indx-1,"score"]))
#       #cat("this is delta(t)",delta_t,"\n")
#       for (i in 1:nopts) {
#         action <- as.character(colnames(Att)[i])
#         #cat("this is current indexed action",action,"\n")
#         #cat("this is previous human action",h_act_prev,"\n")
#         # Attraction vector update rule
#         Att[state_indx,i] <- (Phi_t*N*Att[state_indx,i]  + ( delta_t[[i]] + (1- delta_t[[i]])*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (Phi_t*N + 1)
#       #Update the value of N 
#       N <- Phi_t*N + 1
#       # Assume human chooses action probabilistically using softmax on Attraction values
#       probs <- exp(lambda*Att[new_state,])/sum(exp(lambda*Att[new_state,]))
#       # Get actual human action and compute likelihood
#       h_act <- as.character(game_data[indx,"h_action"])
#       act_index <- match(h_act, colnames(Att))
#       lik_hum[t] <- probs[[act_index]]
#       curr_state <- new_state
#       # if (indx == 50) {
#       #   cat("this is the attraction matrix at last round of", "\n")
#       #   print(Att)
#       # }

### Second attempt at ST_EWA with states and no initial guess of initial states. 
```{r}
ST_EWA_STATES <- function(par,data,return_value){

  lambda <- par[1]

@@

    if(data[t,"round"] == 1) {
      Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
      state_vec <- switch(as.character(t_game), rps=states_RPS, fwg = states_FWG, numbers = states_NUM)
      reward <- switch(as.character(t_game), rps=reward_RPS, fwg = reward_FWG, numbers = reward_NUM)
      game_data <- subset(data, game == t_game)

      # first round is uniform prediction
      lik_hum[t] <- 1/nopts
      # Asumme N(0) = 1 for now...see discussion in paper
      N <- 1.0
      # initialize H_t by looking at first action 
      # H_t <- rep(0.0,nopts)
      a_act <- as.character(game_data[1,"a_action"])
      H_t <- as.numeric(colnames(Att) == a_act)

      # Rounds 2 to end 
    } else {

      indx <- data[t,"round"]

      # Get reward and past human action
      h_act_prev <- as.character(game_data[indx-1,"h_action"])
      a_act_prev <- as.character(game_data[indx-1,"a_action"])
      curr_state <- paste0(h_act_prev,a_act_prev)

      h_act <- as.character(game_data[indx,"h_action"])
      a_act <- as.character(game_data[indx,"a_action"])
      new_state <- paste0(h_act,a_act)

      # Assume human chooses action probabilistically using softmax on Attraction values
      probs <- exp(lambda*Att[curr_state,])/sum(exp(lambda*Att[curr_state,]))

      # Get actual human action and compute likelihood
      act_index <- match(h_act, colnames(Att))
      lik_hum[t] <- probs[[act_index]]

      # Update rule:
      # Estimate R(t) (recent history) then H(t) (history) then phi(t) (change detector)
      R_t <- as.numeric(colnames(Att) == a_act)
      #cat("This is R(t)",R_t,"\n")

      H_t <- (H_t*(indx-1) + R_t)/(indx)
      #cat("this is H(t)",H_t,"\n")

@@

      # Estimate vector Delta(t)
      delta_t <- as.numeric(reward[,a_act] >= as.numeric(game_data[indx,"score"]))
      #cat("this is delta(t)",delta_t,"\n")

      # This is a vectorised ST_EWA update rule, easier to follow the loop below in comments
      state_indx <- match(curr_state,state_vec)
      Att[state_indx,] <- (Phi_t*N*Att[state_indx,]  + ( delta_t + (1- delta_t)*(colnames(Att) == h_act))*as.numeric(reward[colnames(Att),a_act])) / (Phi_t*N + 1)

      ##############
      # for (i in 1:nopts) {
      #   action <- as.character(colnames(Att)[i])
      #   state_indx <- match(curr_state,state_vec)
      #   #cat("this is the current state index", state_indx, "\n")
      # 
      #   # Attraction vector update rule
      #   Att[state_indx,i] <- (Phi_t*N*Att[state_indx,i]  + ( delta_t[[i]] + (1- delta_t[[i]])*(action == h_act))*as.numeric(reward[action,a_act])) / (Phi_t*N + 1)
      # 
      # }
      # 
      #cat("this is the current Att vector", Att, "\n")
      ###################

@@

      #Update the state 
      curr_state <- new_state 
    }
   }  
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)

@@

```{r}
# data12 = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# ST_EWA_STATES(0.5, data12, "-2loglik")
# optim(10.0,fn=ST_EWA_STATES,gr = NULL, data=data12,"-2loglik", lower = 0.0, upper = 200, method="L-BFGS-B")

```{r, cache=TRUE}

ST_EWA_STATES_modelling <- list()
for(id in unique(dat$human_id)) {
  ST_EWA_STATES_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
#   ST_EWA_STATES_modelling[[id]] <- optim(10.0,fn=ST_EWA_STATES,gr = NULL, data=tdat,"-2loglik", lower = 0.0, upper = 200, method="L-BFGS-B")
# }
  ST_EWA_STATES_modelling[[id]] <- DEoptim(fn=ST_EWA_STATES, lower = 0.0, upper = 100.0, data=tdat, "-2loglik", control=list(trace = FALSE,parallelType=1))
}

save(ST_EWA_STATES_modelling,file="ST_EWA_STATES_modelling.RData")

@@

## Putting results together 
```{r}
# load("QL_modelling.RData")
# load("QL_states_modelling.RData")
# load("EWA_modelling.RData")
# load("EWA_self_modelling.RData")
# load("MBM_modelling.RData")
# load("ST_EWA_STATES_modelling.RData")

@@

                         # Self-Tuning EWA (only 1 parameter)
                          # Parametric EWA BIC
                         # "EWA_self_2LL" = EWA_self_modelling[[id]]$optim$bestval,
                         # "EWA_self_BIC" = EWA_self_modelling[[id]]$optim$bestval + 1*log(150),
                         # #Lambda is a parameter of the softmax choice function (inverse Temperature)
                         # "EWA_self_Lambda" = EWA_self_modelling[[id]]$optim$bestmem[1],

                         # self_tuning EWA with states

                         "ST_EWA_STATES_BIC" = ST_EWA_STATES_modelling[[id]]$optim$bestval + 1*log(150),
                         "ST_EWA_STATES_Lambda" = ST_EWA_STATES_modelling[[id]]$optim$bestmem[1]

@@

comp_BICs <- All_results[c("ID","Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","ST_EWA_STATES_BIC")]

@@

# Table_results <- table(All_results[, "condition"],c("random","ToM_BT","ToM_NBT","QL", "QL_states","EWA","S_EWA","ST_EWA_STATES")[apply(All_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC","ST_EWA_STATES_BIC")],1,which.min)])

# We choose to omit the simple QL and EWA and opt to only include the ones with states as they are more relevant.

Table_results <- table(All_results[, "condition"],c("random","ToM_BT","ToM_NBT", "QL_states","ST_EWA_STATES")[apply(All_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_states_BIC","ST_EWA_STATES_BIC")],1,which.min)])

@@

exp1_model_comp <- data.frame()
for(id in unique(dat$human_id)) {
  tdat <- subset(dat,human_id == id)
  tot_score <- sum(tdat$score)
  tot_time <- sum(tdat$human_rt)
  #early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg"))
  early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg" | game =="numbers") )
  tr_score <- sum(early_dat$score)
  id_results <- subset(All_results, ID == id)
  min_BIC <- apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_states_BIC","ST_EWA_STATES_BIC")],1,min)

  best_model <- c("random","ToM_Game_Tr","ToM_No_Tr", "QL_states","ST_EWA_STATES")[apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_states_BIC","ST_EWA_STATES_BIC")],1,which.min)]

  exp1_model_comp <- rbind(exp1_model_comp,
                       data.frame(
                         "human_id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         "Early_game_score" = tr_score,
                         "Total_score" = tot_score,
                         "Best_model" = best_model,
                         "Total_time" = sum(tdat$human_rt),
                         "TR_minus_NT_BIC" = id_results[,"Bayes_Tr_BIC"] - id_results[,"Bayes_No_Tr_BIC"],
                         "Rand_minus_best_BIC" =  id_results[,"Random_BIC"] - min_BIC

@@

# More exploratory work......
